### 2021-10-08 docs/paper/sections/4_implementation.tex
Revise paper and apply feedback by @scheibel
- Restructure sections
- Shorten things
- Rephrasing
- Improve typesetting

In our implementation, we confine ourselves to analyzing node.js projects written in JavaScript.
This decision is based on the large prevalence of JavaScript\footnote{GitHut 2.0: A small place to discover languages in GitHub. \url{https://madnight.github.io/githut/\#/pushes}} as well as the first-rate extent of the npm ecosystem\footnote{Modulecounts: \url{http://www.modulecounts.com/}} which has been unsurpassed since 2015.
In the following, we describe how we apply the single steps of the approach presented above to this particular domain.

\subsection{Downstream dependency collection}
\label{sec:implementation/dependency_collection}

The de facto standard package repository for node.js is npm.\footnote{\url{https://www.npmjs.com/}}
To fetch downstream dependency packages from the npm registry, we use the npm package npm-dependants\footnote{\url{https://www.npmjs.com/package/npm-dependants}} which scrapes the dependents list from npmjs.com as there does not exist a publicly available API for these data.
To keep the footprint of our implementation small, we only scrape a part of the dependents list.
We then request basic metadata for each found package from the npm registry, in particular including a URL to the source code archive of the latest package version (commonly referred to as \emph{tarball}), by using the package-metadata package.\footnote{\url{https://www.npmjs.com/package/package-metadata}}
Finally, we download and extract each source code archive from the npm registry with the help of the download-package-tarball package.\footnote{\url{https://www.npmjs.com/package/download-package-tarball}}

\begin{lstlisting}[language=json,
	label=sec:implementation/package.json,
	caption={An example \code{package.json} manifest file for the \code{ripple-lib} package specifying multiple dependencies (truncated).},
	float=tpb,
	floatplacement=tbp,
	abovecaptionskip=-5pt
]
{
	"name": "ripple-lib",
	"version": "1.10.0",
	"description": "An API for the XRP Ledger",
	"dependencies": {
		"bignumber.js": "^9.0.0",
		"https-proxy-agent": "^5.0.0",
		"jsonschema": "1.2.2"
	}
}
\end{lstlisting}

For the second approach, using a code search engine, we have identified Sourcegraph\footnote{\url{https://sourcegraph.com/}} as a promising platform.
npm packages specify their dependencies in a package.json metadata file (see \cref{sec:implementation/package.json}).
To find all packages depending on a package with the name \code{<pkg>}, we can search all available package.json files for the substring \code{"<pkg>":}.
Refined with some additional constraints to filter out package copies from \code{node\_modules} folders that downstream developers have pushed accidentally, and to limit the number of results retrieved, our final Sourcegraph query has the following form:

{
	\lstset{frame=none,
		aboveskip=3mm,
		belowskip=3mm,
		xleftmargin=3mm,
		xrightmargin=3mm
	}
	\begin{lstlisting}[language=sourcegraph]
"<pkg>":
	file:package\.json
	-file:node_modules/
	select:file
	count:<limit>
	\end{lstlisting}
}

We submit this search query by sending it to the publicly available Sourcegraph GraphQL API\footnote{\url{https://docs.sourcegraph.com/api/graphql}} by using the npm package graphql-request\footnote{\url{https://www.npmjs.com/package/graphql-request}}.
For each hit, we request additional information from the API about the matched repository, including the name of the depending package and the URL to the original repository.
After retrieving a search result, we download a snapshot of the full repository by using the download-git-repo package\footnote{\url{https://www.npmjs.com/package/download-git-repo}} which unifies access to the APIs of different code host providers such as GitHub or BitBucket.

Additionally, we enrich the metadata of each package with a few metrics indicating its popularity from the associated GitHub repository if specified.
The GitHub API is accessed via the official octokit library.\footnote{\url{https://octokit.github.io/}}

\subsection{Usage sample mining}
\label{sec:implementation/usage_mining}

For parsing the dependencies, we utilize the public TypeScript Compiler API.\footnote{\url{https://github.com/Microsoft/TypeScript/wiki/Using-the-Compiler-API}}
TypeScript is an extension of JavaScript that adds strong typing to the former; typing is based on a combination of implicit type inference and explicit type declarations.
As the type-checking capabilities of the TypeScript compiler are optional and can be turned off via configuration, we consider the JavaScript syntax a subset of TypeScript syntax.
As a side benefit, our implementation is also able to analyze packages implemented in true TypeScript.

The TypeScript compiler's \emph{binder} component annotates every node of the parsed program with a type symbol if it can resolve one.
After performing a type analysis of the dependency by instrumenting the binder, we collect all nodes from the type-annotated ASF that refer to the target package by traversing it with a visitor function.
This visitor function matches each node against a list of defined patterns (similar to \cref{fig:usage_mining/patterns}) based on the syntax kinds of the node, its ancestors, and its descendants, and then looks up the declaration of the node's type symbol to compare it with the source directory of the target package.

Normally, foreign types can only be resolved if they are declared in a dependency module that has been installed into the \code{node\_modules} folder of the repository.
Because downloading and installing all dependency modules for every downstream repository of the target package would increase the computational resources of the usage mining drastically and eventually violate \cref{req1}, we skip this step but instead modify the module resolution strategy of the TypeScript compiler host.
To do so, we extend the \code{paths} parameter from the default compiler options\footnote{\url{https://www.typescriptlang.org/tsconfig\#paths}} and insert a new entry that remaps accesses to the target package name to the separate source code directory of the target package.

\begin{figure*}[b]
	\centering
	\begin{subfigure}{.32\linewidth}
		\includegraphics[width=\linewidth]{sections/6_implementation/extension/dependencies.png}
		\caption[LoF entry]{The dependency browser.}
		\label{fig:implementation/presentation/screenshot/dependency_browser}
	\end{subfigure}
	\hfill
	\begin{subfigure}{.32\linewidth}
		\includegraphics[width=\linewidth]{sections/6_implementation/extension/references.png}
		\caption[LoF entry]{The usage browser.}
		\label{fig:implementation/presentation/screenshot/usage_browser}
	\end{subfigure}
	\hfill
	\begin{subfigure}{.32\linewidth}
		\includegraphics[width=\linewidth]{sections/6_implementation/extension/codelens.png}
		\caption[LoF entry]{The CodeLens integration for the class \code{GraphQLError}.}
		\label{fig:implementation/presentation/screenshot/codelens}
	\end{subfigure}

	\caption{Screenshots of our Visual Studio Code extension used to explore downstream dependencies of the npm package graphql.}
	\label{fig:implementation/presentation/screenshots}
\end{figure*}

\subsection{Presentation of results}
\label{sec:implementation/presentation}

After all references have been collected, a proper UI is still required to provide users easy access to these data that fulfills the requirements mentioned above.
To satisfy \cref{req2} and make our tool available in the usual working environment of users, we implement it as an extension to a popular IDE.
One of the most popular IDEs that support JavaScript, and the one with the highest annual growth, is Visual Studio Code.\footnote{\url{https://code.visualstudio.com/}}\footnote{Top IDE Index by Pierre Carbonnelle (2021): \url{https://pypl.github.io/IDE.html}}
As it also provides a comprehensive set of APIs for extension developers, we have decided to roll our tool out as a VS Code extension.\footnote{To support other researchers in reusing our solution, we also provide a CLI for our implementation.}

To support developers in answering the questions raised in \cref{sec:framework}, we have designed three key views for our extension (see \cref{fig:implementation/presentation/screenshots}):

\begin{enumerate}[label=(\roman*)]
	\item The \emph{dependency browser} allows to explore all downstream dependencies and, grouped for each dependency, all their references to the target package.
	\item The \emph{usage browser} displays all public package members and, grouped for each member, all dependencies and their references to this member.
	\item The \emph{CodeLens integration} provides quick access to a slice of the usage browser that is attached to the definition of each package member right in the source code editor.
\end{enumerate}

@@

We implement each of these views in the VS Code Extension API (VSCE)\footnote{\url{https://code.visualstudio.com/api}}, we use the Tree View API and the CodeLens API.
One challenge has been to deliver fast results and not to block the user interface because even inspite of the lightweight mining approaches we chose to satisfy \cref{req1} above, querying, downloading, and analyzing each dependency takes a few seconds (see \cref{sec:evaluation}).
To overcome this issue, we use pagination wherever possible and apply the observer pattern to push incremental UI updates.
As the VSCE API per se does not provide for multithreading or multiprocessing operations but suggests a Promise/A+-driven dataflow, we adopt this style for our object model by using the JavaScript concepts asynchronous functions, \code{AsyncIterator}s, and \code{Promise.all} to avoid busy waiting inside the data collection process.

>>>>

In our implementation, we confine ourselves to analyzing JavaScript projects for node.js.
This decision is based on the large prevalence of JavaScript\footnote{GitHut 2.0: A small place to discover languages in GitHub. \url{https://madnight.github.io/githut/\#/pushes}} as well as the first-rate extent of the npm ecosystem which has been unsurpassed since 2015\footnote{Modulecounts: \url{http://www.modulecounts.com/}}.
In the following, we describe how we apply the single steps of the approach presented above to this domain.

\subsection{Downstream Dependency Collection}
\label{sec:implementation/dependency_collection}

The de facto standard package repository for node.js is npm.
To fetch downstream dependency packages from the npm registry, we use the npm package \code{npm-dependants}\footnote{All packages can be retrieved from \url{npmjs.com}} which scrapes the dependents list from npmjs.com as there does not exist a publicly available API for these data.
To keep the footprint of our implementation small, we only scrape a part of the dependents list.
We then request basic metadata for each found package from the npm registry, including the \emph{tarball} URL of the latest package version, using the package \code{package-metadata}.
Finally, we download and extract each tarball using the package \code{download-package-tarball}.

\begin{lstlisting}[language=json,
	label=sec:implementation/package.json,
	caption={An example \code{package.json} file for the \code{ripple-lib} package specifying multiple dependencies (truncated).},
	float=tpb,
	floatplacement=tbp,
	abovecaptionskip=-5pt
]
{
	"name": "ripple-lib",
	"version": "1.10.0",
	"description": "An API for the XRP Ledger",
	"dependencies": {
		"bignumber.js": "^9.0.0",
		"https-proxy-agent": "^5.0.0",
		"jsonschema": "1.2.2"
	}
}
\end{lstlisting}

For the second approach based on a code search engine, we have identified Sourcegraph\footnote{\url{https://sourcegraph.com/}} as a promising platform.
npm packages specify their dependencies in a package.json metadata file (see \cref{sec:implementation/package.json}).
To find all packages depending on a certain package, we can search all available \code{package.json} files for a reference to this package identifier.
We refine our search query with some additional constraints to filter out package copies from \code{node_modules} folders that downstream developers have pushed accidentally, and to limit the number of results retrieved.
We submit this search query to the publicly available GraphQL\footnote{\url{https://graphql.com/}} API of Sourcegraph\footnote{\url{https://docs.sourcegraph.com/api/graphql}} by using the npm package \code{graphql-request}.
After retrieving a search result, we download a snapshot of the full repository by using the package \code{download-git-repo}.

Additionally, we enrich the metadata of each package with a few metrics indicating its popularity from the associated GitHub repository if specified.
The GitHub API is accessed via the official \emph{octokit} library\footnote{\url{https://octokit.github.io/}}.

\subsection{Usage Sample Mining}
\label{sec:implementation/usage_mining}

For parsing the dependencies, we utilize the \emph{TypeScript Compiler API}\footnote{\url{https://github.com/Microsoft/TypeScript/wiki/Using-the-Compiler-API}}.
TypeScript is an extension of JavaScript that adds strong typing to the former; typing is based on a combination of implicit type inference and explicit type declarations.
As the type-checking capabilities of the TypeScript compiler are optional and can be turned off via configuration, we consider the JavaScript syntax a subset of TypeScript syntax.
As a side benefit, our implementation is also able to analyze packages implemented in true TypeScript.

The TypeScript compiler's \emph{binder} component annotates every node of the parsed program with a type symbol if it can resolve one.
After performing a type analysis of the dependency by instrumenting the binder, we collect all nodes from the type-annotated ASF that refer to the target package by traversing it with a visitor function.
This visitor function matches each node against a list of defined patterns (similar to \cref{fig:approach/usage_mining/patterns}) based on the syntax kinds of the node, its ancestors, and its descendants, and then performs a lookup of the node's type symbol's declaration to compare it with the source directory of the target package.

Normally, foreign types can only be resolved if they are declared in a dependency module that has been installed into the \code{node_modules} folder of the repository.
Because downloading and installing all dependency modules for every downstream repository of the target package would increase the computational resources of the usage mining drastically and eventually violate \cref{req1}, we skip this step but instead modify the module resolution strategy of the TypeScript compiler host.
To do so, we extend the \code{paths} parameter from the default compiler options\footnote{\url{https://www.typescriptlang.org/tsconfig\#paths}} and insert a new entry that remaps accesses to the target package name to the separate source code directory of the target package.

\begin{figure*}
	\centering
	\begin{subfigure}{.32\linewidth}
		\includegraphics[width=\linewidth]{sections/4_implementation/extension/dependencies.png}
		\caption[LoF entry]{The dependency browser.}
		\label{fig:implementation/presentation/screenshot/dependency_browser}
	\end{subfigure}
	\hfill
	\begin{subfigure}{.32\linewidth}
		\includegraphics[width=\linewidth]{sections/4_implementation/extension/references.png}
		\caption[LoF entry]{The usage browser.}
		\label{fig:implementation/presentation/screenshot/usage_browser}
	\end{subfigure}
	\hfill
	\begin{subfigure}{.32\linewidth}
		\includegraphics[width=\linewidth]{sections/4_implementation/extension/codelens.png}
		\caption[LoF entry]{The CodeLens integration for the class \code{GraphQLError}.}
		\label{fig:implementation/presentation/screenshot/codelens}
	\end{subfigure}

	\caption{Screenshots of our VS Code extension used to explore downstream dependencies of the npm package graphql.}
	\label{fig:implementation/presentation/screenshots}
\end{figure*}

\subsection{Presentation of Results}
\label{sec:implementation/presentation}

After all references have been collected, a proper UI is still required to provide users easy access to these data that fulfills the requirements mentioned above.
To satisfy \cref{req2} and make our tool available in the usual working environment of users, we implement it as an extension to the \emph{Visual Studio Code} IDE\footnote{\url{https://code.visualstudio.com/}}.
VS Code is one of the most popular IDEs that support JavaScript, and the one with the highest annual growth.\footnote{Top IDE Index by Pierre Carbonnelle (2021): \url{https://pypl.github.io/IDE.html}}
It also provides a comprehensive set of APIs for extension developers.
To support other researchers in reusing our solution, we also provide a CLI.

To support developers in answering the questions raised in \cref{sec:approach}, we have designed three key views for our extension (see \cref{fig:implementation/presentation/screenshots}):

\begin{enumerate}[label=(\roman*)]
	\item The \emph{dependency browser} allows to explore all downstream dependencies and, grouped for each dependency, all their references to the target package.
	\item The \emph{usage browser} displays all public package members and, grouped for each member, all dependencies and their references to this member.
	\item The \emph{CodeLens integration} provides quick access to a slice usage samples and is attached to the definition of each package member in the source code editor.
\end{enumerate}

@@

We implement each of these views in the VS Code Extension API (VSCE)\footnote{\url{https://code.visualstudio.com/api}}, we use the \emph{Tree View API} and the \emph{CodeLens API}.
One challenge has been to deliver fast results and not to block the user interface because even in spite of the lightweight mining approaches we chose to satisfy \cref{req1} above, querying, downloading, and analyzing each dependency takes a few seconds (see \cref{sec:evaluation}).
To overcome this issue, we use pagination wherever possible and apply the observer pattern to push incremental UI updates.
As the VSCE API per se does not provide for multithreading or multiprocessing operations but suggests a Promise/A+-driven dataflow, we adopt this style for our object model by using the JavaScript concepts asynchronous functions, \code{AsyncIterator}s, and \code{Promise.all} to avoid busy waiting inside the data collection process.

### 2021-10-08 docs/paper/sections/5_evaluation.tex
Revise paper and apply feedback by @scheibel
- Restructure sections
- Shorten things
- Rephrasing
- Improve typesetting

In the following, we are going to investigate each of these questions.

@@

To assess the quality of dependency collection, we analyze it with respect to extent, precision, recall, and performance.
\Cref{tab:evaluation/resqu1/quantities} shows the number of dependencies that were collected from each data source following the proposed approaches for a set of arbitrarily selected npm packages.
For each package, we have annotated a subset of the collected dependencies (max. 20 dependencies per package) manually to identify and classify false positive hits.

\begin{table}
	\centering
	\small
	\input{sections/7_evaluation/resqu1/quantities}
	\caption{Quantity and false-positive rates (FPR) of downstream dependencies found by both approaches for selected packages.}
	\label{tab:evaluation/resqu1/quantities}
\end{table}

\subsubsection{Extent}
\label{sec:evaluation/resqu1/extent}

Regarding the total number of dependencies collected by each approach, no clear trend in favor of any approach can be ascertained for small or medium packages (having less than 10,000 stars on GitHub).
For larger packages (having at least 10,000 stars), however, the number of dependencies collected from the npm registry stagnates near to 400 hits; beyond this limit, the npm registry will return internal server errors only.
For Sourcegraph, on the other hand, we have not hit any limitations so far; currently, they do not provide any official documentation for the exact rate limits.

@@

\subsubsection{Precision}
\label{sec:evaluation/resqu1/precision}

The precision of collected dependencies is drastically lower for dependencies found on Sourcegraph than for such found on npm.
\Cref{fig:evaluation/resqu1/fp_causes} collates the causes we have identified for these false positives; most frequently, repositories specify a dependency on the target package in their package manifest file but do not actually import this package at any place.

\begin{figure}
	\centering
	\small
	\input{sections/7_evaluation/resqu1/fp_causes}
	\caption{Causes for false-positive dependency matches and their frequencies.}
	\label{fig:evaluation/resqu1/fp_causes}
\end{figure}

In some situations, this may happen if the package is a plugin for another package or if it is invoked as a CLI from a build script (\TODO{C'mon, we could improve data classification for this}), but in the majority of repositories, developers trivially appear to have specified the upstream dependency by accident (e.g., while copying a \code{package.json} file over from another package), or to have forgotten to remove the upstream dependency after switching away from using the package.
Rarer causes include repositories that only add type definitions for TypeScript to a package but do not actually import it, as well as, especially for repositories found on Sourcegraph, other packages required by a repository declaring a peer dependency on the target package which needs to be fulfilled by the depending repository.

@@

\subsubsection{Recall}
\label{sec:evaluation/resqu1/recall}

Besides the precision of collected dependencies, their recall is also of interest.
As the dependencies are collected from two giant datasets in extracts only, a quantitative analysis of false-negative hits based on manual annotation of these datasets would be too expensive for this paper.
Nevertheless, there are some causes that will prevent a dependency from being found by our approaches:

\begin{itemize}
	\item Packages without a proper dependency manifest cannot be detected by our approaches which are based on exactly these metadata.
	\item On npm, naturally packages only will be found, creating a bias towards rather generic and professional software.
	\item On both platforms, results will be sorted based on intransparent criteria (which, as we however speculate, include the number of direct dependents on npm and the recent update frequency on Sourcegraph, resp.).
		As we only fetch the first dependencies from both data sources, this sorting is a likely source of further biases.
		These biases could be fought by always fetching all dependencies, but this would drastically reduce the overall performance.
\end{itemize}

\subsubsection{Performance}
\label{sec:evaluation/resqu1/performance}

\begin{table}
	\centering
	\small
	\begin{threeparttable}
		\input{sections/7_evaluation/resqu1/performance}
		\begin{tablenotes}
			\footnotesize
			\item[\alphtnotetext{1}] Hardware specifications of the test machine: 7 vCPUs Intel Xeon Cascade Lake @ \si{2.80}{GHz}, internet downspeed ~1.8 \si{Gbit}/\si{s}.
			\item[\alphtnotetext{2}] Effective speed while downloading multiple packages in parallel.
		\end{tablenotes}
	\end{threeparttable}

	\caption{Key performance metrics for both dependency collection methods.}
	\label{tab:evaluation/resqu1/performance}
\end{table}

@@

\subsubsection{Precision}
\label{sec:evaluation/resqu2/precision}

Under laboratory conditions, false-positive usage samples cannot be emitted by our approach if we assume the correctness of the TypeScript compiler.
There are only two exceptions to this invariant:
\begin{enumerate*}[label=(\roman*)]
	\item If there is a name collision between two packages, invalid or false positives usage samples may be emitted.
		However, the npm infrastructure attempts to rule out these collisions by using the package name as a unique ID for every published package.
	\item If dependency developers deliberately interfere with the TypeScript compiler (for instance, by suppressing type errors using a \verb|@ts-ignore| comment\footnote{\url{https://www.typescriptlang.org/docs/handbook/release-notes/typescript-2-6.html\#suppress-errors-in-ts-files-using--ts-ignore-comments}}), both false positive and false negative matches are possible.
\end{enumerate*}

\subsubsection{Recall}
\label{sec:evaluation/resqu2/recall}

@@

\begin{itemize}
	\item Many projects on OSS platforms have complex build configurations that include additional transpilation or code generation steps before a final version of the source code is reached that is valid to the node.js interpreter or the TypeScript compiler.
		Unless we add explicit support for such build configurations, type analysis and thus usage mining for these projects will fail.
	\item In some situations, additional type definitions are required to perform a complete type inference of a dependency.
		This is the case for every parametrized callback from a package for that type definitions are not available.
		This limitation could be resolved by downloading or generating the type definitions for all upstream dependencies of every downstream dependency; this, on the other hand, would reduce the performance of the approach.
	\item In general, the static type analyzer of TypeScript has its limitations.
		For instance, the type analysis will have a limited recall for certain control flow patterns or metaprogramming constructs such as meta-circular evaluation using the the \code{eval()} function or dynamic function binding using \code{Function.bind()} and friends.
\end{itemize}

\subsubsection{Performance}
\label{sec:evaluation/resqu2/performance}

The total time spent for performing type analysis for a repository from the experimental dataset and mining usage samples averages ca. \num{3} seconds.
The complete ASF of an average parsed repository consumes between \SI{10}{MB} and \SI{500}{MB} of RAM, depending on the size of the analyzed repository.

@@

To assess the applicability of our proposed tool, we investigate the different options of users who would like to answer the questions raised in \cref{sec:framework}.
We further analyze how far our tool meets the requirements posed above.

\subsubsection{Answering user questions}

In order to answer \cref{q1} without our tool, it would be possible to access the aforementioned data sources manually.
Both npm and Sourcegraph provide suitable web interfaces for surveying dependent packages or search results, resp.
To understand the intent of a dependency, users can read its documentation online, if any, or browse its implementation.
The source code of repositories can be browsed on Sourcegraph, but on npm, this is not possible, and users need to either manually download and extract the package or follow the link to an open-source code repository if the package provides one.

@@

In order to answer \cref{q2,q3} without our tool, users need to scan each dependency found in the previous step separately to count or read all usage samples, resp.
If they are intending to analyze the usage of a member with an unambiguous name, they can perform an expeditious string search.
For repositories found on Sourcegraph, this is either possible by starting a new search on the platform or by searching the source code in the original code host.
If the source code is not available online, it needs to be downloaded first and searched locally.
If the member of interest has a confusable name, a string search could not lead to the desired result.
In this case, users need to download and install the project and view it in an IDE that supports reference search (for instance, there is a built-in feature for this in VS Code).
Alternatively, they can try to use the precise code intelligence mode on Sourcegraph (see \cref{sec:related_work/usage_samples}).
However, as of today, precise code intelligence is only available for a small number of repositories as both the dependent repository and the dependency repository are required to provide a special index file.

Within our tool, all usage samples can be collected automatically and be merged into a single list.
Once the collected dependencies have been processed, users can select a member of interest and view all its usage samples with a single click.
For the usage analysis of Java APIs, there is also the tool \caps{Exapus}\citep{de2013multi} which differentiates between different kinds of member usages such as instantiation of vs. inheritance of a class.
Such a classification mechanism could also improve the usability of our tool

\subsubsection{Meeting requirements}

We break down \cref{req1} into three acceptance criteria: application liveness, small resource footprint, and easy setup.

As a result of the numbers mentioned in the evaluation of the previous research questions, our tool is able to process 5 up to 12 dependencies per minute while requiring about \SI{500}{MB} RAM in total and less than \SI{30}{MB} ROM per package.
I.e., from the moment the tool is activated, the first results usually take less than 10 seconds to appear on the UI.
This is an acceptable delay with regard to Shneiderman's definition of acceptable application response times, so the liveness criterion is fulfilled.\footnote{Ben Shneiderman and Catherine Plaisant. 2010. Designing the user interface: Strategies for effective human-computer interaction. Pearson Education India.}
Even if the tool is run for 10 minutes, it will occupy less than \SI{4}{GB} of ROM that can be released at any time, which we consider a small resource footprint at a time where modern operating systems require at least \SI{64}{GB} ROM; thus, also the second acceptance criterion is fulfilled.
As our extension can be installed with two clicks from the Visual Studio Code Extension Marketplace, it is also easy to set up and thus fulfills the first requirement.

\Cref{req2} effectively calls for a minimum number of context switches that are required to use the tool, i.e., for minimizing the temporal, spatial, and semantic distance\footnote{David Ungar, Henry Lieberman, and Christopher Fry. 1997. Debugging and the experience of immediacy. Commun. ACM 40, 4 (1997), 38–43.} between the downstream dependency UI and the IDE used by package developers.

Once the data are available in the UI, interaction with them is possible with the same speed as with local source files in the package.
This is compatible with Shneiderman's requirement to frequent tasks and indicates a low temporal distance.

The spatial distance varies for the different views of the tool:
For the dependency browser and the usage browser, a separate toolbar has to be opened, which indicates a higher spatial distance; still, the artifacts are available in the same IDE.
On the other hand, the dependency artifacts were not displayed in the IDE without our tool and have no strong relationship to the already existing artifacts, so any closer integration into an existing view would increase unintended coupling between independent domains.
The code annotations are very close to the original source code and thus have a small spatial distance.

The semantic distance can be described as the perceived similarity of artifacts displayed both in the tool's UI and in the existing IDE.
Such artifacts are the package and dependency members displayed in the usage browser of our tool.
Most items in the usage browser bear the same label as the corresponding identifier in the source code, which reduces the semantic distance.
However, some nodes in the path of dependency members correspond to anonymous expressions in the source code that do not bear a label themselves and thus are labeled using a generic fallback token in our tool (see \cref{fig:implementation/presentation/screenshot/dependency_browser}).
This does not match the labeling convention in VS Code (as displayed in the outline view of the explorer activity bar), where a target identifier is determined from the context of the expression.
For such nodes, the current semantic distance of our tool is still higher than necessary.

>>>>

In the following, we will investigate each question.

@@

To assess the quality of dependency collection, we analyze it with respect to extent, precision, recall, and performance.
\Cref{tab:evaluation/resqu1/quantities} shows the number of dependencies that we collected to build an experimental dataset from each data source following the proposed approaches for a set of arbitrarily selected npm packages.
For each package, we have annotated a subset of the collected dependencies (max. 20 dependencies per package) to identify and classify false positive hits.

\begin{table}
	\caption{Quantity and false-positive rates (FPR) of downstream dependencies found by both approaches for selected packages.}
	\label{tab:evaluation/resqu1/quantities}

	\centering
	\small
	\input{sections/5_evaluation/resqu1/quantities}
\end{table}

\paragraph{Extent}
\label{sec:evaluation/resqu1/extent}

Regarding the total number of dependencies collected by each approach, no clear trend in favor of any approach can be ascertained for small or medium packages (having less than \num{10,000} stars on GitHub).
For larger packages (having at least \num{10,000} stars), however, the number of dependencies collected from the npm registry stagnates near to 400 hits; beyond this limit, the npm registry will return internal server errors only.
For Sourcegraph, we have not hit any limitations so far; currently, they do not provide any official documentation for the exact rate limits.

@@

\paragraph{Precision}
\label{sec:evaluation/resqu1/precision}

The precision of collected dependencies is drastically lower for dependencies found on Sourcegraph than for such found on npm.
\Cref{fig:evaluation/resqu1/fp_causes} collates the causes we have identified for false positives; most frequently, repositories specify a dependency on a target package in their package manifest file but do not import this package at any place.

\begin{figure}
	\centering
	\small
	% Fix font size of \thead
	\renewcommand\theadfont{\small}
	\input{sections/5_evaluation/resqu1/fp_causes}

	\caption{Causes for false-positive dependency matches and their frequencies.}
	\label{fig:evaluation/resqu1/fp_causes}
\end{figure}

In some situations, this may happen if the package is a plugin for another package or if it is invoked as a CLI from a build script, but in the majority of repositories, developers trivially appear to have specified the upstream dependency by accident (e.g., while copying a \code{package.json} file over from another package), or to have forgotten to remove the upstream dependency after switching away from using the package.
Rarer causes include repositories that only add type definitions for TypeScript to a package but do not actually import it, as well as, especially for repositories found on Sourcegraph, other packages required by a repository declaring a peer dependency on the target package which needs to be fulfilled by the depending repository.

@@

\paragraph{Recall}
\label{sec:evaluation/resqu1/recall}

Besides the precision of collected dependencies, their recall is also of interest.
As the dependencies are collected from two giant datasets in excerpts only, a quantitative analysis of false-negative hits based on manual annotation of these datasets would be too expensive for this paper.
Nevertheless, there are some causes that will prevent a dependency from being found by our approaches:

\begin{itemize}
	\item Packages without a proper dependency manifest cannot be detected by our approaches which are based on exactly these metadata.
	\item On npm, only published packages will be found, creating a bias towards generic and professional software.
	\item On both platforms, results will be sorted based on intransparent criteria (which, as we speculate, include the number of direct dependents on npm and the recent update frequency on Sourcegraph, resp.).
		As we only fetch the first dependencies from both data sources, this is a likely source of further biases.
		These could be fought by always fetching all dependencies, but this would drastically reduce the overall performance.
\end{itemize}

\paragraph{Performance}
\label{sec:evaluation/resqu1/performance}

\begin{table}
	\caption{Key performance metrics for both dependency collection methods.}
	\label{tab:evaluation/resqu1/performance}

	\centering
	\begin{threeparttable}
		% Fix font size of \thead
		\renewcommand\theadfont{\normalsize}
		\input{sections/5_evaluation/resqu1/performance}
		\begin{tablenotes}
			\footnotesize
			\item[\alphtnotetext{1}] Hardware specifications of the test machine: 7 vCPUs Intel Xeon Cascade Lake @ \SI{2.80}{\giga\hertz}, internet downspeed ~\SI{1.8}{\giga\bit/\second}.
			\item[\alphtnotetext{2}] Effective speed while downloading multiple packages in parallel.
		\end{tablenotes}
	\end{threeparttable}
\end{table}

@@

\paragraph{Precision}
\label{sec:evaluation/resqu2/precision}

Under laboratory conditions, false-positive usage samples cannot be emitted by our approach if we assume the correctness of the TypeScript compiler.
There are only two exceptions to this invariant:
\begin{enumerate*}[label=(\roman*)]
	\item In case of a name collision between two packages, invalid or false positives usage samples may be emitted.
		However, the npm infrastructure attempts to rule out these collisions by using the package name as a unique ID for every published package.
	\item If dependency developers deliberately interfere with the TypeScript compiler (for instance, by suppressing type errors using a \verb|@ts-ignore| comment\footnote{\url{https://www.typescriptlang.org/docs/handbook/release-notes/typescript-2-6.html\#suppress-errors-in-ts-files-using--ts-ignore-comments}}), both false positive and false negative matches are possible.
\end{enumerate*}

Additionally, some npm packages contain minified sources only that are still searchable valid code but can provide a significantly reduced readability.

\paragraph{Recall}
\label{sec:evaluation/resqu2/recall}

@@

\begin{itemize}
	\item Many projects on OSS platforms have complex build configurations that include additional transpilation or code generation steps before a final version of the source code is reached that is valid to the node.js interpreter or the TypeScript compiler.
		Unless we add explicit support for such build configurations, type analysis and usage mining for these projects will fail.
	\item In some situations, additional type definitions are required to perform a complete type inference of a dependency.
		This applies to every parametrized callback from a package for that type definitions are not available.
		This limitation could be resolved by downloading or generating type definitions for all upstream dependencies of every downstream dependency; this, however, would reduce the performance of the approach.
	\item In general, the static type analyzer of TypeScript has its limitations.
		For instance, the type analysis will have a limited recall for certain control flow patterns or metaprogramming constructs such as meta-circular evaluation using the the \code{eval()} function or dynamic function binding using \code{Function.bind()} etc.
\end{itemize}

\paragraph{Performance}
\label{sec:evaluation/resqu2/performance}

The total time spent performing type analysis for a repository from the experimental dataset and mining usage samples averages ca. \num{3} seconds.
The complete ASF of an average parsed repository consumes between \SI{10}{\mega\byte} and \SI{500}{\mega\byte} RAM, depending on the size of the repository.

@@

To assess the applicability of our proposed tool, we investigate the different options of users who would like to answer the questions raised in \cref{sec:approach}.
We further analyze how far our tool meets the requirements posed above.

\paragraph{Answering User Questions}

To answer \cref{q1} without our tool, users could access the aforementioned data sources manually to search for and view downstream dependencies.
The source code of repositories can be browsed on Sourcegraph, but if an npm package is not available on an OSS platform, users will need to download and extract it.

@@

To answer \cref{q2,q3} without our tool, users need to scan each found dependency separately to count or read all usage samples.
If they are intending to analyze the usage of a member with an unambiguous name, they can perform an expeditious string search.
This is possible using Sourcegraph if the dependency is indexed there; otherwise, users will need to download the repository and search it locally.
If the member of interest has an ambiguous name, users will need to download the project and view it in an IDE such as VS Code that supports reference search.
Alternatively, they can try the precise code intelligence mode on Sourcegraph (see \cref{sec:related_work/usage_samples}); however, this feature is available for a small number of repositories only.

Within our tool, all usage samples can be collected automatically and be merged into a single list.
Once the collected dependencies have been processed, users can select a member of interest and view all its usage samples with a single click.
For the usage analysis of Java APIs, the tool \emph{Exapus} \citep{de2013multi} also considers different kinds of member usages such as instantiation vs. inheritance of a class; a similar classification could also improve the usability of our tool.

\paragraph{Meeting Requirements}

We break down \cref{req1} into three acceptance criteria: application liveness, a small resource footprint, and an easy setup.

As learned from the previous research questions, our tool is able to process 5 up to 12 dependencies per minute while requiring about \SI{500}{\mega\byte} RAM in total and less than \SI{30}{\mega\byte} ROM per package.
After the tool is activated, the first results usually take less than 10 seconds to appear on the UI.
This is an acceptable delay regarding Shneiderman's definition of acceptable application response times \citep{shneiderman2010designing}, so the liveness criterion is fulfilled.
Even if the tool downloads a few hundred dependencies, it will occupy less than \SI{10}{\giga\byte} ROM that can be released at any time, which we consider a small resource footprint at a time where modern operating systems require \SI{64}{\giga\byte} ROM; thus, also the second acceptance criterion is fulfilled.
As our extension can be installed with two clicks from the Visual Studio Code Extension Marketplace, it also fulfills the setup criterion and thus the first requirement.

\Cref{req2} effectively calls for a minimum number of context switches that are required to use the tool, i.e., for minimizing the temporal, spatial, and semantic distance \citep{ungar1997debugging} between the downstream dependency UI and the IDE used by package developers.

Once the data are available in the UI, interaction with them is possible with the same speed as with local source files in the package.
This meets Shneiderman's requirement to frequent tasks and indicates a low temporal distance.

The spatial distance varies for the different views of the tool:
For the dependency browser and the usage browser, a separate toolbar has to be opened, which indicates a higher spatial distance; still, the artifacts are available in the same IDE.
Also, the dependency artifacts have no strong relationship to any existing IDE artifacts, so any closer integration into an existing view would increase coupling between independent domains.
The code annotations are disaplyed close to the original source code with a small spatial distance.

The semantic distance can be described as the perceived similarity of artifacts displayed both in the tool's UI and in the existing IDE, i.e., the package and dependency members.
Most items in the usage browser bear the same label as the corresponding identifier in the source code, which reduces the semantic distance.
However, dependency member path nodes that refer to anonymous source code expressions are displayed differently than in the outline view of VS Code (see \cref{fig:implementation/presentation/screenshot/dependency_browser}), increasing the semantic distance a bit.
